---
title: "R Notebook"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

REFERENCES:
http://stackoverflow.com/questions/1330989/rotating-and-spacing-axis-labels-in-ggplot2
http://stackoverflow.com/questions/30057765/histogram-ggplot-show-count-label-for-each-bin-for-each-category

---
```{r}
library(ggplot2)
```

# Functions

The following function is the function which will be used to get the mode of the data

```{r}
mymode <- function(x){
  xtable <- table(x)
  idx <- xtable == max(xtable)
  names(xtable)[idx]
}
```

# Read in the data from the csv file

```{r}
churn_data <- read.csv("./eurocomPHONEchurners.csv")
```

First thing to do is to print out the data summary.

```{r}
summary(churn_data)
```

# Preprocess the data


## Identify duplicate values
ref: http://www.cookbook-r.com/Manipulating_data/Finding_and_removing_duplicate_records/

```{r}
churn_data[duplicated(churn_data$CUST_ID), ]
```

## Remove the duplicates
```{r}
churn_data <- churn_data[!duplicated(churn_data$CUST_ID), ]
```


```{r}
nrow(churn_data)
```


## Identify null values 

The summary data can be used to identify the columns which contain null values. 

The columns with null values are:

* MINUTES_3MONTHS_AGO (3 records)
* CUST_MOS (3 records)
* TOT_MINUTES_USAGE (4 records)
* PHONE_PLAN (4 records)
* EDUCATION (8 records) 

## Addressing the issues of the missing values. 

For the numeric values they will be replaced with median value for the columns

### MINUTES_3MONTHS_AGO
```{r}
median(churn_data$MINUTES_3MONTHS_AGO, na.rm = TRUE)
```

```{r}
MINUTES_3MONTHS_AGO_median <- median(churn_data$MINUTES_3MONTHS_AGO, na.rm = TRUE)
```

```{r}
churn_data$MINUTES_3MONTHS_AGO[is.na(churn_data$MINUTES_3MONTHS_AGO)] <- MINUTES_3MONTHS_AGO_median 
```


### CUST_MOS
```{r}
median(churn_data$CUST_MOS, na.rm = TRUE)
```

```{r}
CUST_MOS_median <- median(churn_data$CUST_MOS, na.rm = TRUE)
```

```{r}
churn_data$CUST_MOS[is.na(churn_data$CUST_MOS)] <- CUST_MOS_median 
```

### TOT_MINUTES_USAGE
```{r}
median(churn_data$TOT_MINUTES_USAGE, na.rm = TRUE)
```

```{r}
TOT_MINUTES_USAGE_median <- median(churn_data$TOT_MINUTES_USAGE, na.rm = TRUE)
```

```{r}
churn_data$TOT_MINUTES_USAGE[is.na(churn_data$TOT_MINUTES_USAGE)] <- TOT_MINUTES_USAGE_median 
```

### PHONE_PLAN

The first thing to do was to ceate a data frame which is comprised of the gender and phone plan

```{r}

  phone_plan_gender_subset <- data.frame(churn_data$PHONE_PLAN, churn_data$GENDER)

```

The data frame is then filtered again so it just contains data for record for male records

```{r}
  phone_plan_male_subset <- phone_plan_gender_subset[phone_plan_gender_subset$churn_data.GENDER == 'M', 1]
```

```{r}
  phone_plan_male_subset[1:10]
```

The mode of the phone plan is stored formale records

```{r}

  print(phone_plan_male_mode <- mymode(phone_plan_male_subset))

```

The data frame is then filtered again so it just contains data for record for female records

```{r}
  phone_plan_female_subset <- phone_plan_gender_subset[phone_plan_gender_subset$churn_data.GENDER == 'F', 1]
```

```{r}
  phone_plan_female_subset[1:10]
```


```{r}
  print(phone_plan_female_mode <- mymode(phone_plan_female_subset))
```


We then find the indexes of the phone plan records with empty values

```{r}
  no_phone_plans <- which(churn_data$PHONE_PLAN == "")
```


```{r}
for (index in no_phone_plans)
{
  if (churn_data$GENDER[index] == 'M') 
  {
    churn_data$PHONE_PLAN[index] <- phone_plan_male_mode
  } 
  else 
  {
    churn_data$PHONE_PLAN[index] <- phone_plan_female_mode
  }
}

```

```{r}
  churn_data$PHONE_PLAN[no_phone_plans]
```

```{r}
churn_data$PHONE_PLAN <- droplevels(churn_data$PHONE_PLAN)
```

```{r}
summary(churn_data$PHONE_PLAN)
```

### EDUCATION
The first thing to do was to ceate a data frame which is comprised of the gender and education

```{r}

  education_gender_subset <- data.frame(churn_data$EDUCATION, churn_data$GENDER)

```

The data frame is then filtered again so it just contains data for record for male records

```{r}
  no_education_male_subset <- education_gender_subset[education_gender_subset$churn_data.GENDER == 'M', 1]
```


The mode of the phone plan is stored formale records

```{r}
  no_education_male_mode <- mymode(no_education_male_subset)
```

```{r}
  no_education_male_mode
```

The data frame is then filtered again so it just contains data for record for female records

```{r}
  no_education_female_subset <- education_gender_subset[education_gender_subset$churn_data.GENDER == 'F', 1]
```


```{r}

  no_education_female_mode <- mymode(no_education_female_subset)

```

```{r}

  no_education_female_mode

```


We then find the indexes of the phone plan records with empty values

```{r}
  no_education <- which(churn_data$EDUCATION == "")
```

```{r}
for (index in no_education)
{
  if (churn_data$GENDER[index] == 'M') 
  {
    churn_data$EDUCATION[index] <- no_education_male_mode
  } 
  else 
  {
    churn_data$EDUCATION[index] <- no_education_female_mode
  }
}
```

```{r}
churn_data$EDUCATION[no_education]
```

```{r}
churn_data$EDUCATION <- droplevels(churn_data$EDUCATION)
```

```{r}
summary(churn_data)
```


```{r}
churn_data$INCOME <- cut(churn_data$INCOME, breaks=c(0,38000,88000,Inf), right = FALSE, labels=c("low income","medium income","high income"))
```

# Predictor Information

## CUST_ID

Unique Account Number that identifies an account holder

+-------------------+-------------------------------------------+
| Predictor         | CUST_ID
+===================+===========================================+
| attribute type    | Nominal                                 ` |
+-------------------+-------------------------------------------+
| Max               | `r max(churn_data$CUST_ID)`               |
+-------------------+-------------------------------------------+
| Min               | `r min(churn_data$CUST_ID)`               |
+-------------------+-------------------------------------------+


## AREA_CODE

Geographical area account holder resides

+-------------------+-------------------------------------------+
| Predictor         | AREA_CODE
+===================+===========================================+
| attribute type    | Nominal                                   |
+-------------------+-------------------------------------------+
| %Missing Values   | 0%                                        |
+-------------------+-------------------------------------------+
| Mode              | `r mymode(churn_data$AREA_CODE)`          |
+-------------------+-------------------------------------------+


```{r}
plot_data <- data.frame(table(churn_data$AREA_CODE, churn_data$CHURNER))
colnames(plot_data) <- c("Area_Code", "Churned", "Count" )
ggplot(data=plot_data, aes(x = Area_Code, y = Count, fill = Churned, label = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(size = 3, position = position_stack(vjust = 0.5))
```


## LONGDIST_FLAG

Whether has signed up for the off-peak long distance call package 

+-------------------+-------------------------------------------+
| Predictor         | LONGDIST_FLAG
+===================+===========================================+
| attribute type    | Nominal                       |
+-------------------+-------------------------------------------+
| %Missing Values   |                     0                     |
+-------------------+-------------------------------------------+
| Mode              | `r mymode(churn_data$LONGDIST_FLAG)`|
+-------------------+-------------------------------------------+

```{r}
plot_data <- data.frame(table(churn_data$LONGDIST_FLAG, churn_data$CHURNER))
colnames(plot_data) <- c("LONGDIST_FLAG", "Churned", "Count" )
ggplot(data=plot_data, aes(x = LONGDIST_FLAG, y = Count, fill = Churned, label = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(size = 3, position = position_stack(vjust = 0.5))
```

## CALLWAITING_FLAG

Whether the customer has call waiting

+-------------------+-------------------------------------------+
| Predictor         | CALLWAITING_FLAG
+===================+===========================================+
| attribute type    | Nominal                                   |
+-------------------+-------------------------------------------+
| %Missing Values   | 0%                                        |
+-------------------+-------------------------------------------+
| Mode              | `r mymode(churn_data$CALLWAITING_FLAG)` |
+-------------------+-------------------------------------------+

```{r}
plot_data <- data.frame(table(churn_data$CALLWAITING_FLAG, churn_data$CHURNER))
colnames(plot_data) <- c("CALLWAITING_FLAG", "Churned", "Count")
ggplot(data=plot_data, aes(x = CALLWAITING_FLAG, y = Count, fill = Churned, label = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(size = 3, position = position_stack(vjust = 0.5))
```

## VOICEMAIL_FLAG

Whether the customer has voice mail

+-------------------+-------------------------------------------+
| Predictor         | VOICEMAIL_FLAG
+===================+===========================================+
| attribute type    | Nominal                              |
+-------------------+-------------------------------------------+
| %Missing Values   | 0%                                        |
+-------------------+-------------------------------------------+
| Mode              | `r mymode(churn_data$VOICEMAIL_FLAG)` |
+-------------------+-------------------------------------------+

```{r}
plot_data <- data.frame(table(churn_data$VOICEMAIL_FLAG, churn_data$CHURNER))
colnames(plot_data) <- c("VOICEMAIL_FLAG", "Churned", "Count" )
ggplot(data=plot_data, aes(x = VOICEMAIL_FLAG, y = Count, fill = Churned, label = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(size = 3, position = position_stack(vjust = 0.5))
```

## MOBILE_PLAN

Has the customer signed up to the mobile phone plan

+-------------------+-------------------------------------------+
| Predictor         | MOBILE_PLAN
+===================+===========================================+
| attribute type    | Nominal                                   |
+-------------------+-------------------------------------------+
| %Missing Values   | 0%                                        |
+-------------------+-------------------------------------------+
| Mode              | `r mymode(churn_data$MOBILE_PLAN)` |
+-------------------+-------------------------------------------+

```{r}
plot_data <- data.frame(table(churn_data$MOBILE_PLAN, churn_data$CHURNER))
colnames(plot_data) <- c("MOBILE_PLAN", "Churned", "Count" )
ggplot(data=plot_data, aes(x = MOBILE_PLAN, y = Count, fill = Churned, label = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(size = 3, position = position_stack(vjust = 0.5))
```

## CONVERGENT_BILLING

All service charges consolidated onto one bill

+-------------------+-------------------------------------------+
| Predictor         | CONVERGENT_BILLING
+===================+===========================================+
| attribute type    | Nominal                                   |
+-------------------+-------------------------------------------+
| %Missing Values   | 0%                                        |
+-------------------+-------------------------------------------+
| Mode              | `r mymode(churn_data$CONVERGENT_BILLING)` |
+-------------------+-------------------------------------------+

```{r}
plot_data <- data.frame(table(churn_data$CONVERGENT_BILLING, churn_data$CHURNER))
colnames(plot_data) <- c("CONVERGENT_BILLING", "Churned", "Count" )
ggplot(data=plot_data, aes(x = CONVERGENT_BILLING, y = Count, fill = Churned, label = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(size = 3, position = position_stack(vjust = 0.5))
```

## GENDER

Account holder's gender

+-------------------+-------------------------------------------+
| Predictor         | GENDER
+===================+===========================================+
| attribute type    | Nominal                                   |
+-------------------+-------------------------------------------+
| %Missing Values   | 0%                                        |
+-------------------+-------------------------------------------+
| Mode              | `r mymode(churn_data$GENDER)` |
+-------------------+-------------------------------------------+

```{r}
plot_data <- data.frame(table(churn_data$GENDER, churn_data$CHURNER))
colnames(plot_data) <- c("GENDER", "Churned", "Count" )
ggplot(data=plot_data, aes(x = GENDER, y = Count, fill = Churned, label = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(size = 3, position = position_stack(vjust = 0.5))
```

## INCOME

Account holder's annual income (€)

+-------------------+-------------------------------------------+
| Predictor         | INCOME
+===================+===========================================+
| attribute type    | Ordinal                                   |
+-------------------+-------------------------------------------+
| %Missing Values   | 0%                                        |
+-------------------+-------------------------------------------+
| Mode              | `r mymode(churn_data$INCOME)` |
+-------------------+-------------------------------------------+

```{r}
plot_data <- data.frame(table(churn_data$INCOME, churn_data$CHURNER))
colnames(plot_data) <- c("INCOME", "Churned", "Count" )
ggplot(data=plot_data, aes(x = INCOME, y = Count, fill = Churned, label = Count)) + 
  geom_histogram(stat = "identity") + 
  geom_text(size = 3, position = position_stack(vjust = 0.5))
```

## PHONE_PLAN

The phone plan the customer has signed up for national, euro-zone, international (outside Euro-zone) and promo_plan (signed up to the promotional plan)

+-------------------+-------------------------------------------+
| Predictor         | PHONE_PLAN
+===================+===========================================+
| attribute type    | Nominal                                   |
+-------------------+-------------------------------------------+
| %Missing Values   | 0%                                        |
+-------------------+-------------------------------------------+
| Mode              | `r mymode(churn_data$PHONE_PLAN)` |
+-------------------+-------------------------------------------+

```{r}
plot_data <- data.frame(table(churn_data$PHONE_PLAN, churn_data$CHURNER))
colnames(plot_data) <- c("PHONE_PLAN", "Churned", "Count" )
ggplot(data=plot_data, aes(x = PHONE_PLAN, y = Count, fill = Churned, label = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(size = 3, position = position_stack(vjust = 0.5))
```

## EDUCATION

Highest Level of education attainment the account holder has achieved

+-------------------+-------------------------------------------+
| Predictor         | EDUCATION
+===================+===========================================+
| attribute type    | Nominal                                   |
+-------------------+-------------------------------------------+
| %Missing Values   | 0%                                        |
+-------------------+-------------------------------------------+
| Mode              | `r mymode(churn_data$EDUCATION)` |
+-------------------+-------------------------------------------+


```{r}
plot_data <- data.frame(table(churn_data$EDUCATION, churn_data$CHURNER))
colnames(plot_data) <- c("EDUCATION", "Churned", "Count" )
ggplot(data=plot_data, aes(x = EDUCATION, y = Count, fill = Churned, label = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(size = 3, position = position_stack(vjust = 0.5))
```

## MINUTES_CURR_MONTH

Phone Minutes currently for current months (to the time the data was extracted)

+-------------------+-------------------------------------------+
| Predictor         | MINUTES_CURR_MONTH
+===================+===========================================+
| attribute type    | Numeric                                   |
+-------------------+-------------------------------------------+
| %Missing Values   | 0%                                        |
+-------------------+-------------------------------------------+
| Max               | `r max(churn_data$MINUTES_CURR_MONTH)`    |
+-------------------+-------------------------------------------+
| Min               | `r min(churn_data$MINUTES_CURR_MONTH)`    |
+-------------------+-------------------------------------------+
| Mean              | `r mean(churn_data$MINUTES_CURR_MONTH)`   |
+-------------------+-------------------------------------------+
| Mode              | `r mymode(churn_data$MINUTES_CURR_MONTH)` |
+-------------------+-------------------------------------------+
| median            | `r median(churn_data$MINUTES_CURR_MONTH)` |
+-------------------+-------------------------------------------+
| Standard deviation| `r sd(churn_data$MINUTES_CURR_MONTH)`     |
+-------------------+-------------------------------------------+

### MINUTES_CURR_MONTH Skewness

```{r}
(3 * (mean(churn_data$MINUTES_PREV_MONTH) - median(churn_data$MINUTES_PREV_MONTH)))/ sd(churn_data$MINUTES_PREV_MONTH)
```


```{r}
plot_data <- data.frame(table(churn_data$MINUTES_CURR_MONTH, churn_data$CHURNER))
colnames(plot_data) <- c("MINUTES_CURR_MONTH", "CHURNER", "COUNT")
ggplot(data=plot_data, aes(x = MINUTES_CURR_MONTH, y=COUNT, fill = CHURNER, label=COUNT)) +
  geom_histogram(stat="identity", width = 1) +
  geom_text(size = 2, position = position_stack(vjust = 0.5))+ 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## MINUTES_PREV_MONTH

Phone Minutes used in the previous month

+-------------------+-------------------------------------------+
| Predictor         | MINUTES_PREV_MONTH
+===================+===========================================+
| attribute type    | Numerical                       |
+-------------------+-------------------------------------------+
| %Missing Values   |                                     |
+-------------------+-------------------------------------------+
| Max               | `r max(churn_data$MINUTES_PREV_MONTH)`    |
+-------------------+-------------------------------------------+
| Min               | `r min(churn_data$MINUTES_PREV_MONTH)`    |
+-------------------+-------------------------------------------+
| Mean              | `r mean(churn_data$MINUTES_PREV_MONTH)`   |
+-------------------+-------------------------------------------+
| Mode              | `r mymode(churn_data$MINUTES_PREV_MONTH)` |
+-------------------+-------------------------------------------+
| median            | `r median(churn_data$MINUTES_PREV_MONTH)` |
+-------------------+-------------------------------------------+

### MINUTES_PREV_MONTH Skewness

```{r}
(3 * (mean(churn_data$MINUTES_PREV_MONTH) - median(churn_data$MINUTES_PREV_MONTH)))/ sd(churn_data$MINUTES_PREV_MONTH)
```

```{r}
plot_data <- data.frame(table(churn_data$MINUTES_PREV_MONTH, churn_data$CHURNER))
colnames(plot_data) <- c("MINUTES_PREV_MONTH", "CHURNER", "COUNT")
ggplot(data=plot_data, aes(x = MINUTES_PREV_MONTH, y=COUNT, fill = CHURNER, label=COUNT)) +
  geom_histogram(stat="identity", width = 1) +
  geom_text(size = 2, position = position_stack(vjust = 0.5))+ 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
## MINUTES_3MONTHS_AGO

Phone Minutes used in the 3 month PREVIOUS 

+-------------------+-------------------------------------------+
| Predictor         | MINUTES_3MONTHS_AGO
+===================+===========================================+
| attribute type    | Numerical                       |
+-------------------+-------------------------------------------+
| %Missing Values   |                                          |
+-------------------+-------------------------------------------+
| Max               | `r max(churn_data$MINUTES_3MONTHS_AGO)`   |
+-------------------+-------------------------------------------+
| Min               | `r min(churn_data$MINUTES_3MONTHS_AGO)`   |
+-------------------+-------------------------------------------+
| Mean              | `r mean(churn_data$MINUTES_3MONTHS_AGO)`  |
+-------------------+-------------------------------------------+
| Mode              | `r mymode(churn_data$MINUTES_3MONTHS_AGO)`|
+-------------------+-------------------------------------------+
| median            | `r median(churn_data$MINUTES_3MONTHS_AGO)`|
+-------------------+-------------------------------------------+

### MINUTES_3MONTHS_AGO Skewness

Right-skewness data – Is positive, as mean is greater than the median
```{r}
(3 * (mean(churn_data$MINUTES_3MONTHS_AGO) - median(churn_data$MINUTES_3MONTHS_AGO)))/ sd(churn_data$MINUTES_3MONTHS_AGO)
```

```{r}
plot_data <- data.frame(table(churn_data$MINUTES_3MONTHS_AGO, churn_data$CHURNER))
colnames(plot_data) <- c("MINUTES_3MONTHS_AGO", "CHURNER", "COUNT")
ggplot(data=plot_data, aes(x = MINUTES_3MONTHS_AGO, y=COUNT, fill = CHURNER, label=COUNT)) +
  geom_histogram(stat="identity", width = 1) +
  geom_text(size = 2, position = position_stack(vjust = 0.5))+ 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## CUST_MOS

The number of continuous months the Customer is with the provider

+-------------------+-------------------------------------------+
| Predictor         | CUST_MOS
+===================+===========================================+
| attribute type    | Numerical                      |
+-------------------+-------------------------------------------+
| %Missing Values   |                     0                     |
+-------------------+-------------------------------------------+
| Max               | `r max(churn_data$CUST_MOS)`   |
+-------------------+-------------------------------------------+
| Min               | `r min(churn_data$CUST_MOS)`   |
+-------------------+-------------------------------------------+
| Mean              | `r mean(churn_data$CUST_MOS)`  |
+-------------------+-------------------------------------------+
| Mode              | `r mymode(churn_data$CUST_MOS)`|
+-------------------+-------------------------------------------+
| median            | `r median(churn_data$CUST_MOS)`|
+-------------------+-------------------------------------------+

### CUST_MOS Skewness

Right-skewness data – Is positive, as mean is greater than the median
```{r}
(3 * (mean(churn_data$CUST_MOS) - median(churn_data$CUST_MOS)))/ sd(churn_data$CUST_MOS)
```


```{r}
plot_data <- data.frame(table(churn_data$CUST_MOS, churn_data$CHURNER))
colnames(plot_data) <- c("CUST_MOS", "CHURNER", "COUNT")
ggplot(data=plot_data, aes(x = CUST_MOS, y=COUNT, fill = CHURNER, label=COUNT)) +
  geom_histogram(stat="identity", width = 1) +
  geom_text(size = 2, angle = 90, position = position_stack(vjust = 0.5))+ 
  theme(axis.text.x = element_text(size=6, angle = 90, hjust = .5))
``` 

## TOT_MINUTES_USAGE

The total number of minutes used to date

+-------------------+-------------------------------------------+-------------------+-------------------------------------------+
| Predictor         | TOT_MINUTES_USAGE                         | Predictor         | TOT_MINUTES_USAGE
+===================+===========================================+===================+===========================================+
| attribute type    | Numeric                                   |
+-------------------+-------------------------------------------+
| %Missing Values   | 0%                                        |
+-------------------+-------------------------------------------+
| Max               | `r max(churn_data$TOT_MINUTES_USAGE)`    |
+-------------------+-------------------------------------------+
| Min               | `r min(churn_data$TOT_MINUTES_USAGE)`    |
+-------------------+-------------------------------------------+
| Mean              | `r mean(churn_data$TOT_MINUTES_USAGE)`   |
+-------------------+-------------------------------------------+
| Mode              | `r mymode(churn_data$TOT_MINUTES_USAGE)` |
+-------------------+-------------------------------------------+
| median            | `r median(churn_data$TOT_MINUTES_USAGE)` |
+-------------------+-------------------------------------------+
| Standard deviation| `r sd(churn_data$TOT_MINUTES_USAGE)`     |
+-------------------+-------------------------------------------+


### TOT_MINUTES_USAGE Skewness

Right-skewness data – Is positive, as mean is greater than the median
```{r}
(3 * (mean(churn_data$TOT_MINUTES_USAGE) - median(churn_data$TOT_MINUTES_USAGE)))/ sd(churn_data$TOT_MINUTES_USAGE)
```


```{r}
plot_data <- data.frame(table(churn_data$TOT_MINUTES_USAGE, churn_data$CHURNER))
colnames(plot_data) <- c("TOT_MINUTES_USAGE", "CHURNER", "COUNT")
ggplot(data=plot_data, aes(x = TOT_MINUTES_USAGE, y=COUNT, fill = CHURNER, label=COUNT)) +
  geom_histogram(stat="identity", width = 1) +
  geom_text(size = 2, angle = 90, position = position_stack(vjust = 0.5))+ 
  theme(axis.text.x = element_text(size=6, angle = 90, hjust = .5))
``` 


## NUM_LINES

The number of fixed lines the customer has leased.

+-------------------+-------------------------------------------+
| Predictor         | NUM_LINES
+===================+===========================================+
| attribute type    | Numeric                                   |
+-------------------+-------------------------------------------+
| %Missing Values   | 0%                                        |
+-------------------+-------------------------------------------+
| Max               | `r max(churn_data$NUM_LINES)`    |
+-------------------+-------------------------------------------+
| Min               | `r min(churn_data$NUM_LINES)`    |
+-------------------+-------------------------------------------+
| Mean              | `r mean(churn_data$NUM_LINES)`   |
+-------------------+-------------------------------------------+
| Mode              | `r mymode(churn_data$NUM_LINES)` |
+-------------------+-------------------------------------------+
| median            | `r median(churn_data$NUM_LINES)` |
+-------------------+-------------------------------------------+
| Standard deviation| `r sd(churn_data$NUM_LINES)`     |
+-------------------+-------------------------------------------+

### NUM_LINES Skewness

Right-skewness data – Is positive, as mean is greater than the median
```{r}
(3 * (mean(churn_data$NUM_LINES) - median(churn_data$NUM_LINES)))/ sd(churn_data$NUM_LINES)
```


```{r}
plot_data <- data.frame(table(churn_data$NUM_LINES, churn_data$CHURNER))
colnames(plot_data) <- c("NUM_LINES", "CHURNER", "COUNT")
ggplot(data=plot_data, aes(x = NUM_LINES, y=COUNT, fill = CHURNER, label=COUNT)) +
  geom_histogram(stat="identity", width = 1) +
  geom_text(size = 2, position = position_stack(vjust = 0.5))+ 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

# Skewness for numeric data

Right-skewness data – Is positive, as mean is greater than the median
Left skewness data – Mean is smaller than the median, generating negative values
Perfectly symmetric data – mean, median and mode are equal, so skewness is zero


```{r}
(3 * (mean(churn_data$MINUTES_3MONTHS_AGO) - median(churn_data$MINUTES_3MONTHS_AGO)))/ sd(churn_data$MINUTES_3MONTHS_AGO)
```

```{r}
(3 * (mean(churn_data$MINUTES_PREV_MONTH) - median(churn_data$MINUTES_PREV_MONTH)))/ sd(churn_data$MINUTES_PREV_MONTH)
```

```{r}
(3 * (mean(churn_data$MINUTES_CURR_MONTH) - median(churn_data$MINUTES_CURR_MONTH)))/ sd(churn_data$MINUTES_CURR_MONTH)
```

```{r}
(3 * (mean(churn_data$TOT_MINUTES_USAGE) - median(churn_data$TOT_MINUTES_USAGE)))/ sd(churn_data$TOT_MINUTES_USAGE)
```

```{r}
(3 * (mean(churn_data$NUM_LINES) - median(churn_data$NUM_LINES)))/ sd(churn_data$NUM_LINES)
```

```{r}
(3 * (mean(churn_data$CUST_MOS) - median(churn_data$CUST_MOS)))/ sd(churn_data$CUST_MOS)
```


# Graphical methods to identify outliers

## MINUTES_3MONTHS_AGO

```{r}
ggplot(data=churn_data, aes(x=CHURNER, y=MINUTES_3MONTHS_AGO))+ geom_boxplot()

```

## MINUTES_PREV_MONTH

```{r}
ggplot(data=churn_data, aes(x=CHURNER, y=MINUTES_PREV_MONTH))+ geom_boxplot()
```

## MINUTES_CURR_MONTH

```{r}
ggplot(data=churn_data, aes(x=CHURNER, y=MINUTES_CURR_MONTH))+ geom_boxplot()
```

## TOT_MINUTES_USAGE

```{r}
ggplot(data=churn_data, aes(x=CHURNER, y=TOT_MINUTES_USAGE))+ geom_boxplot()
```

## NUM_LINES

```{r}
ggplot(data=churn_data, aes(x=CHURNER, y=NUM_LINES))+ geom_boxplot()
```

## CUST_MOS

```{r}
ggplot(data=churn_data, aes(x=CHURNER, y=CUST_MOS))+ geom_boxplot()
```

# Mathematically identifying outliers

Choose a numeric predictor variable that has possible outliers based on your analysis in 2 above. Use the
IQR method and Z-Score Standardisation method to identify outliers. Discuss your findings.

## IQR method

```{r}
minutesQuartileTolerance <- 1.5* IQR(churn_data$TOT_MINUTES_USAGE)
minutesLowerQuantile <- quantile(churn_data$TOT_MINUTES_USAGE, 0.25)
minutesUpperQuantile <- quantile(churn_data$TOT_MINUTES_USAGE, 0.75)

print("lower outliers")
churn_data$TOT_MINUTES_USAGE[churn_data$TOT_MINUTES_USAGE < (minutesLowerQuantile - minutesQuartileTolerance)]

print("upper outliers")
churn_data$TOT_MINUTES_USAGE[churn_data$TOT_MINUTES_USAGE > (minutesUpperQuantile + minutesQuartileTolerance)]

```

## zScore Standardisation method

```{r}
# zScore Standardisation
zscore.TOT_MINUTES_USAGE <-(churn_data$TOT_MINUTES_USAGE - mean(churn_data$TOT_MINUTES_USAGE))/sd(churn_data$TOT_MINUTES_USAGE)
summary(zscore.TOT_MINUTES_USAGE)
```

```{r}
zscore.TOT_MINUTES_USAGE[zscore.TOT_MINUTES_USAGE < -3 | zscore.TOT_MINUTES_USAGE > 3]
```

# Correcting skewness

```{r}
# calculate skewness
TOT_MINUTES_USAGESkewness <- (3 * (mean(churn_data$TOT_MINUTES_USAGE) - median(churn_data$TOT_MINUTES_USAGE)))/ sd(churn_data$TOT_MINUTES_USAGE)
TOT_MINUTES_USAGESkewness
```

## Natural Log Transformation

```{r}
# Natural Log Transformation
natlog.TOT_MINUTES_USAGE <- log(churn_data$TOT_MINUTES_USAGE[churn_data$TOT_MINUTES_USAGE > 0])

# applying log function to skewness
logTOT_MINUTES_USAGESkewness <- (3 * (mean(natlog.TOT_MINUTES_USAGE) - median(natlog.TOT_MINUTES_USAGE)))/ sd(natlog.TOT_MINUTES_USAGE)
logTOT_MINUTES_USAGESkewness
```

## Square Root Transformation

```{r}
# Square Root Transformation
sqrt.TOT_MINUTES_USAGE <- sqrt(churn_data$TOT_MINUTES_USAGE)
# applying sqr root function to skewness

sqrTOT_MINUTES_USAGESkewness <- (3 * (mean(sqrt.TOT_MINUTES_USAGE) - median(sqrt.TOT_MINUTES_USAGE)))/ sd(sqrt.TOT_MINUTES_USAGE)
sqrTOT_MINUTES_USAGESkewness
```

## zScore Standardisation

```{r}
# zScore Standardisation
zscore.TOT_MINUTES_USAGE <-(churn_data$TOT_MINUTES_USAGE - mean(churn_data$TOT_MINUTES_USAGE))/sd(churn_data$TOT_MINUTES_USAGE)

# applying zScore function to skewness
zscore.TOT_MINUTES_USAGE <- (3 * (mean(zscore.TOT_MINUTES_USAGE) - median(zscore.TOT_MINUTES_USAGE)))/ sd(zscore.TOT_MINUTES_USAGE)
zscore.TOT_MINUTES_USAGE
```


# 2D Scatter plots to investigate correlation

2D Scatter plots to investigate correlation between numeric variables

## TOT_MINUTES_USAGE vs CUST_MOS

```{r}
ggplot(churn_data, aes(x=TOT_MINUTES_USAGE, y=CUST_MOS)) +
    geom_point() +
    geom_smooth(method=lm)
```

### Correlation 

```{r}
cor(x=churn_data$TOT_MINUTES_USAGE, y=churn_data$CUST_MOS, use="all.obs", method="pearson")
# http://www.statmethods.net/stats/correlations.html
```

## TOT_MINUTES_USAGE vs MINUTES_CURR_MONTH

```{r}
ggplot(churn_data, aes(x=TOT_MINUTES_USAGE, y=MINUTES_CURR_MONTH)) +
    geom_point() +
    geom_smooth(method=lm)

```

### Correlation 

```{r}
cor(x=churn_data$TOT_MINUTES_USAGE, y=churn_data$MINUTES_CURR_MONTH, use="all.obs", method="pearson")
# http://www.statmethods.net/stats/correlations.html
```

## TOT_MINUTES_USAGE vs MINUTES_PREV_MONTH

```{r}
ggplot(churn_data, aes(x=TOT_MINUTES_USAGE, y=MINUTES_PREV_MONTH)) +
    geom_point() +
    geom_smooth(method=lm)
```

### Correlation 

```{r}
cor(x=churn_data$TOT_MINUTES_USAGE, y=churn_data$MINUTES_PREV_MONTH, use="all.obs", method="pearson")
# http://www.statmethods.net/stats/correlations.html
```

## TOT_MINUTES_USAGE vs MINUTES_3MONTHS_AGO

```{r}
ggplot(churn_data, aes(x=TOT_MINUTES_USAGE, y=MINUTES_3MONTHS_AGO)) +
    geom_point() +
    geom_smooth(method=lm)

```

### Correlation 

```{r}
cor(x=churn_data$TOT_MINUTES_USAGE, y=churn_data$MINUTES_3MONTHS_AGO, use="all.obs", method="pearson")
# http://www.statmethods.net/stats/correlations.html
```

## TOT_MINUTES_USAGE vs NUM_LINES

```{r}
ggplot(churn_data, aes(x=TOT_MINUTES_USAGE, y=NUM_LINES)) +
    geom_point() +
    geom_smooth(method=lm)
```

### Correlation 

```{r}
cor(x=churn_data$TOT_MINUTES_USAGE, y=churn_data$NUM_LINES, use="all.obs", method="pearson")
# http://www.statmethods.net/stats/correlations.html
```

## CUST_MOS vs MINUTES_CURR_MONTH

```{r}
ggplot(churn_data, aes(x=CUST_MOS, y=MINUTES_CURR_MONTH)) +
    geom_point() +
    geom_smooth(method=lm)

```

### Correlation

```{r}
cor(x=churn_data$TOT_MINUTES_USAGE, y=churn_data$MINUTES_CURR_MONTH, use="all.obs", method="pearson")
# http://www.statmethods.net/stats/correlations.html
```

## CUST_MOS vs MINUTES_PREV_MONTH

```{r}
ggplot(churn_data, aes(x=CUST_MOS, y=MINUTES_PREV_MONTH)) +
    geom_point() +
    geom_smooth(method=lm)
```
### Correlation 

```{r}
cor(x=churn_data$TOT_MINUTES_USAGE, y=churn_data$MINUTES_PREV_MONTH, use="all.obs", method="pearson")
# http://www.statmethods.net/stats/correlations.html
```


## CUST_MOS vs MINUTES_3MONTHS_AGO

```{r}
ggplot(churn_data, aes(x=CUST_MOS, y=MINUTES_3MONTHS_AGO)) +
    geom_point() +
    geom_smooth(method=lm)
```

### Correlation 

```{r}
cor(x=churn_data$CUST_MOS, y=churn_data$MINUTES_3MONTHS_AGO, use="all.obs", method="pearson")
# http://www.statmethods.net/stats/correlations.html
```

## CUST_MOS vs NUM_LINES

```{r}
ggplot(churn_data, aes(x=CUST_MOS, y=NUM_LINES)) +
    geom_point() +
    geom_smooth(method=lm)

```

### Correlation 

```{r}
cor(x=churn_data$CUST_MOS, y=churn_data$NUM_LINES, use="all.obs", method="pearson")
# http://www.statmethods.net/stats/correlations.html
```

## MINUTES_CURR_MONTH vs MINUTES_PREV_MONTH

```{r}
ggplot(churn_data, aes(x=MINUTES_CURR_MONTH, y=MINUTES_PREV_MONTH)) +
    geom_point() +
    geom_smooth(method=lm)

```

### Correlation 

```{r}
cor(x=churn_data$MINUTES_CURR_MONTH, y=churn_data$MINUTES_PREV_MONTH, use="all.obs", method="pearson")
# http://www.statmethods.net/stats/correlations.html
```

## MINUTES_CURR_MONTH vs MINUTES_3MONTHS_AGO

```{r}
ggplot(churn_data, aes(x=MINUTES_CURR_MONTH, y=MINUTES_3MONTHS_AGO)) +
    geom_point() +
    geom_smooth(method=lm)

```

### Correlation 

```{r}
cor(x=churn_data$MINUTES_CURR_MONTH, y=churn_data$MINUTES_3MONTHS_AGO, use="all.obs", method="pearson")
# http://www.statmethods.net/stats/correlations.html
```

## MINUTES_CURR_MONTH vs NUM_LINES

```{r}
ggplot(churn_data, aes(x=MINUTES_CURR_MONTH, y=NUM_LINES)) +
    geom_point() +
    geom_smooth(method=lm)

```

### Correlation 

```{r}
cor(x=churn_data$MINUTES_CURR_MONTH, y=churn_data$NUM_LINES, use="all.obs", method="pearson")
# http://www.statmethods.net/stats/correlations.html
```

## MINUTES_PREV_MONTH vs NUM_LINES

```{r}
ggplot(churn_data, aes(x=MINUTES_PREV_MONTH, y=NUM_LINES)) +
    geom_point() +
    geom_smooth(method=lm)

```

### Correlation 

```{r}
cor(x=churn_data$MINUTES_PREV_MONTH, y=churn_data$NUM_LINES, use="all.obs", method="pearson")
# http://www.statmethods.net/stats/correlations.html
```

## MINUTES_PREV_MONTH vs MINUTES_3MONTHS_AGO

```{r}
ggplot(churn_data, aes(x=MINUTES_PREV_MONTH, y=MINUTES_3MONTHS_AGO)) +
    geom_point() +
    geom_smooth(method=lm)

```

### Correlation 

```{r}
cor(x=churn_data$MINUTES_PREV_MONTH, y=churn_data$MINUTES_3MONTHS_AGO, use="all.obs", method="pearson")
# http://www.statmethods.net/stats/correlations.html
```

## MINUTES_3MONTHS_AGO vs NUM_LINES

```{r}
ggplot(churn_data, aes(x=MINUTES_3MONTHS_AGO, y=NUM_LINES)) +
    geom_point() +
    geom_smooth(method=lm)
```

### Correlation 

```{r}
cor(x=churn_data$MINUTES_3MONTHS_AGO, y=churn_data$NUM_LINES, use="all.obs", method="pearson")
# http://www.statmethods.net/stats/correlations.html
```










